Disturbance-aware minimum-time planning strategies for motorsport vehicles with probabilistic safety certificates

### 2025-06-06 intro.tex
[MM] New plot inserted

Following an alternative approach, the largest Lyapunov exponent (LLE) has been proposed as an indicator of local stability~\cite{Meng:AnalysisGlobalCharacteristics:2022}. The LLE quantifies the average exponential rate at which infinitesimal perturbations off a fiducial trajectory grow or decay---thus providing a direct measure of how deviations propagate through the system. 
\textcolor{red}{(negative values denote convergence toward the reference trajectory, while positive values indicate divergence)~\cite{Strogatz:NonlinearDynamicsChaos:2019}. 
%Except in chaotic systems where a positive LLE may still correspond to bounded motion on a strange attractor - a negative LLE typically signals local stability. 
Notably, for trajectories converging to a stable equilibrium, the (negative) LLE asymptotically approaches the real part of the largest eigenvalues of the linearized system at that point. This result has been leveraged in the literature to infer whether a system, not yet at equilibrium, is ultimately converging toward one, thereby enabling a localized assessment of stability based on the transient behavior~\cite{Sadri:StabilityAnalysisNonlinear:2013}.} Applications of this method are particularly relevant to study chaotic vessel motions~\cite{McCue:UseLyapunovExponents:2011} and rotorcraft dynamics~\cite{Tamer:StabilityNonlinearTimeDependent:2016,Cassoni:RotorcraftStabilityAnalysis:2024}, where the primary concern is understanding the long-term evolution of the system and determining its asymptotic behavior.
%However, while Lyapunov-based indicators have been applied in vehicular contexts as well~\cite{Sadri:StabilityAnalysisNonlinear:2013,Meng:AnalysisGlobalCharacteristics:2022}, their relevance here is arguably of reduced interest - not due to their binary nature, but because even identifying asymptotic stability regions offers limited practical insight during planning. A trajectory may lie within the basin of attraction of a distant equilibrium, yet the long-term convergence may be of little operational use. What matters instead is whether, in the short term, a realistically modeled driver - acting in feedback or planning a bit into the future - can ensure that the vehicle's response to perturbations remains within safety-critical constraints, such as grip limits and collision avoidance. This perspective prioritizes short-horizon constraint-satisfying stabilizability over asymptotic notions of stability.
Although Lyapunov-based indicators have been applied in vehicular contexts as well~\cite{Sadri:StabilityAnalysisNonlinear:2013,Meng:AnalysisGlobalCharacteristics:2022}, their utility in planning is limited by the fact that asymptotic stability -- while theoretically informative -- often lacks operational relevance. A point along the planned trajectory may belong to the basin of attraction of a stable equilibrium, but if that equilibrium lies far ahead in time or outside the operational domain (e.g., off track), its practical relevance is questionable.

>>>>

Following an alternative approach, the largest Lyapunov exponent (LLE) has been proposed as an indicator of local stability~\cite{Meng:AnalysisGlobalCharacteristics:2022}. The LLE quantifies the average exponential rate at which infinitesimal perturbations off a fiducial trajectory grow or decay---thus providing a direct measure of how deviations propagate through the system. 
Applications of this method are particularly relevant to study chaotic vessel motions~\cite{McCue:UseLyapunovExponents:2011} and rotorcraft dynamics~\cite{Tamer:StabilityNonlinearTimeDependent:2016,Cassoni:RotorcraftStabilityAnalysis:2024}, where the primary concern is understanding the long-term evolution of the system and determining its asymptotic behavior.
%However, while Lyapunov-based indicators have been applied in vehicular contexts as well~\cite{Sadri:StabilityAnalysisNonlinear:2013,Meng:AnalysisGlobalCharacteristics:2022}, their relevance here is arguably of reduced interest - not due to their binary nature, but because even identifying asymptotic stability regions offers limited practical insight during planning. A trajectory may lie within the basin of attraction of a distant equilibrium, yet the long-term convergence may be of little operational use. What matters instead is whether, in the short term, a realistically modeled driver - acting in feedback or planning a bit into the future - can ensure that the vehicle's response to perturbations remains within safety-critical constraints, such as grip limits and collision avoidance. This perspective prioritizes short-horizon constraint-satisfying stabilizability over asymptotic notions of stability.
Although Lyapunov-based indicators have been applied in vehicular contexts as well~\cite{Sadri:StabilityAnalysisNonlinear:2013,Meng:AnalysisGlobalCharacteristics:2022}, their utility in planning is limited by the fact that asymptotic stability -- while theoretically informative -- often lacks operational relevance. A point along the planned trajectory may belong to the basin of attraction of a stable equilibrium, but if that equilibrium lies far ahead in time or outside the operational domain (e.g., off track), its practical relevance is questionable.

### 2025-06-06 open_loop_planning.tex
[MM] New plot inserted

\ifdefstring{\version}{marco}{
\begin{figure}
	\centering
	\definecolor{mygreen}{RGB}{0,150,0}
	\input{Fig/Tikz_Sources/tikz_covariance_evolution.tex}
	\caption{Cartoon representation (depicted for $H=3$) of the continuity and initialization equations associated with the $H+1$ propagated instances of the covariance matrix introduced in problem~\eqref{eq:DOCPopenloop}. The dashed rectangle highlights the $k$-th discretization step, at which the robust constraints are enforced. Each node corresponds to a specific covariance matrix, where the subscript denotes the current grid index and the superscript indicates the number of propagation steps since initialization. At each step, two key instances are emphasized: the matrix to be initialized at that step (red node), and the matrix that has been propagated over $H$ steps (green node), which is used in the evaluation of the constraint backoff.}
	%\caption{Schematic representation of the continuity and initialization Equations for the $H+1$ versions of the covariance matrix introduced in the discretized OCP. The dashed rectangle indicate the $k$-th step, in which the constraints are formulated. Each node represents a covariance matrix, labeled such that the subscript denotes the current grid step, while the superscript indicates the number of propagation steps it has undergone. At each step, two nodes are highlighted: the covariance matrix to be initialized (red node) and the matrix that has been propagated for $H$ steps (green node), which is used to formulate the robust constraints.}
	\label{fig:DOCPgrid}
\end{figure}
}{}

>>>>

\ifdefstring{\version}{marco}{
\begin{figure}
	\centering
	\definecolor{mygreen}{RGB}{0,150,0}
	\input{Fig/Tikz_Sources/tikz_covariance_evolution.tex}
	\caption{Schematic representation (depicted for $H=3$) of the continuity and initialization equations associated with the $H+1$ propagated instances of the covariance matrix introduced in problem~\eqref{eq:DOCPopenloop}. The dashed rectangle highlights the $k$-th discretization step, at which the robust constraints are enforced. Each node corresponds to a specific covariance matrix, where the subscript denotes the current grid index and the superscript indicates the number of propagation steps since initialization. At each step, two key instances are emphasized: the matrix to be initialized at that step (red node), and the matrix that has been propagated over $H$ steps (green node), which is used in the evaluation of the constraint backoff.}
	%\caption{Schematic representation of the continuity and initialization Equations for the $H+1$ versions of the covariance matrix introduced in the discretized OCP. The dashed rectangle indicate the $k$-th step, in which the constraints are formulated. Each node represents a covariance matrix, labeled such that the subscript denotes the current grid step, while the superscript indicates the number of propagation steps it has undergone. At each step, two nodes are highlighted: the covariance matrix to be initialized (red node) and the matrix that has been propagated for $H$ steps (green node), which is used to formulate the robust constraints.}
	\label{fig:DOCPgrid}
\end{figure}
}{}

### 2025-06-06 performance_comparison.tex
[MM] New plot inserted

\subsection{Robustified constraints comparison}
%It is of interest to analyze how a robustified constraint---or a combination of the two described in Sections~\ref{sec:FLC} and~\ref{sec:TLC}---influences the driving style.
We now investigate how the inclusion of different kinds of robustified constraints --- individually or in combination, as described in Sections~\ref{sec:FLC} and~\ref{sec:TLC} --- influences the driving style.
We compare the nominal feed-forward optimal trajectory, obtained without robustified constraints, with those resulting from a robustified track limit constraint (denoted as TLC), a robustified friction limit constraint (denoted as FLC), and a scenario where both constraints are robustified (denoted as TLC+FLC). All the optimization are obtained with the open-loop method described in Section~\ref{sec:open_loop_planning}, setting $H=4$, and $\ga^\textrm{TLC}=\ga^\textrm{FLC}=1.28$, corresponding to a 90\% probability of satisfying both constraints.

\begin{figure}
	\centering
	\includegraphics{Fig/ol_telemetries.pdf}
\caption{Variations of longitudinal speed $\Delta u$, steering angle $\Delta\delta$, longitudinal force $\Delta F_x$, and lateral deviation $\Delta e$ relative to the nominal trajectory for three schemes --- TLC (track-limit), FLC (friction-limit), and TLC+FLC (both).
Dashed lines in the last three panels mark the sign of the reference signal, indicating whether a positive variation increases or decreases the absolute value (crucial for the turn-dependent variables $\delta$ and $e$ and for distinguishing traction from braking in $F_x$).}

@@

%---Delta key signals

The comparison between the nominal case and the three robustified cases is illustrated in Figure~\ref{fig:ol_telemetries}, in terms of the deviation of selected signals from to their nominal counterparts.
The panels display variations on the longitudinal speed $\De u$ (first panel), wheel steering angle $\De \de$ (second panel), total longitudinal force $\De F_x$ (third panel), and lateral deviation of the center of mass (CoM) from the track centerline $\De e$ (fourth panel).
Except for the $\De u$ panel, each plot includes dashed black lines that indicate the sign of the nominal signal.
These indicators are shown at three distinct vertical levels --- upper, central, and lower --- corresponding respectively to positive, near-zero, and negative values of the nominal signal.

@@

This shift in trajectory and longitudinal speed becomes especially evident when approaching the sharp turn for $\al \in [0.707, 0.712]$. A noticeable increase in longitudinal speed is observed around $\alpha = 0.71$ for the TLC configuration  in the $\De u$ panel. This peak is followed by a sharp drop in speed, indicating intense braking.
The TLC+FLC configuration exhibits a qualitatively similar trend in $\Delta u$, although the peak is significantly lower and the subsequent speed reduction is clearly stronger. This behavior is consistent with the variation of the total longitudinal force $\Delta F_x$ shown in the third panel.
\textcolor{red}{In particular, the TLC and TLC+FLC configurations exhibit a higher longitudinal force while approaching the sharp turn, in the curvilinear abscissa interval $\al\in\left[0.707, 0.712\right]$, resulting in a higher longitudinal speed.
Subsequently, these two configurations apply a lower longitudinal force during braking compared to the nominal case, indicating more intense braking. This behavior explains the sharp reduction in longitudinal speed observed thereafter.}

@@

\begin{figure}
	\centering
	\includegraphics{Fig/ol_saturation.pdf}
	\caption{Axle saturation $S_j(\bx,\bu)$ for front axle (j=1, first panel) and rear axle (j=2, second panel) with respect to the curvilinear abscissa $\al$. The dashed black line corresponds to the reference (nominal) configuration, while the green, red, and blue lines represent the robustified configurations TLC, FLC, and TLC+FLC, respectively.}
	\label{fig:ol_saturation}
\end{figure}

Firstly, we observe that the FLC and TLC+FLC configurations (red and blue lines) never reach full saturation ($S_j=1)$ due to the presence of the back-off term, which enforces a safety margin from the friction limit.
Secondly, we point out how the different trajectories followed by the TLC and TLC+FLC configurations (green and blue lines), compared to the nominal and FLC cases, entail distinct ground force demands, which are clearly visible here in terms of the $S_j$ ratio.
This behavior emerges around the sharp turn and becomes even more evident in the high-speed turn.
Specifically, in the interval $\alpha \in [0.745, 0.762]$, the TLC and TLC+FLC configurations begin to experience high axle loads approximately 12\,m before the nominal trajectory and sustain them for an additional 4.6\,m (TLC, green line) and 9.2\,m (TLC+FLC, blue line) beyond the nominal reference.

@@

\subsection{Open-loop vs closed-loop robust planning}
We now compare the open-loop planning with the closed-loop approach. In particular, we consider open-loop and closed-loop trajectories computed using both the robustified TLC and FLC, with $\ga^\textrm{TLC}=\ga^\textrm{FLC}=1.28$. As before, the open-loop planning is carried out with a prediction horizon of $H=4$ steps.

For the TLC, the back-off term $\be^{TLC}$ in~\eqref{eq:TLC} admits a clear geometrical interpretation: it corresponds to the amount by which the track boundaries are tightened to meet the probabilistic constraint, or, equivalently, to a virtual extension of the vehicle's width. According to the latter interpretation, Figure~\ref{fig:olcl_traj_bands} shows two highlighted bands---red for the open-loop case and cyan for the closed-loop case---corresponding to the area swept by the back-off-extended-vehicle. Ultimately, the TLC enforces precisely that these bands remain entirely within the actual track limits. The comparison between the two regions illustrates how the additional conservativeness of the open-loop approach stems from larger back-off terms, i.e., a wider band that must stay within the track boundaries. The closed-loop method, on the other hand, benefits from the LQR-based feedback, which allows it to maintain smaller back-off terms and thus a narrower band, requiring a reduced safety margin from the track limits.
\begin{figure}
	\centering
	\includegraphics{Fig/olcl_traj_bands.pdf}
	\caption{Lorem ipsum.}
	\label{fig:olcl_traj_bands}
\end{figure}

\subsection{Simulations with noise realization}
Finally, we provide an empirical validation of the closed-loop robustified planning by simulating the evolution of the controlled vehicle with random initial conditions and including noise realizations. The random values for the initial state and the noise follow Gaussian distributions with covariance matrices $\bP_0$ and $\bQ$, respectively, as introduced in~\eqref{eq:dP}. This analysis cannot be performed for the open-loop approach, as it does not compute a stabilizing controller---resulting in trajectories that would significantly deviate from the track limits beyond the prediction horizon.
%For visual clarity, the comparison is limited to trajectories with the robustified TLC, but analogous considerations apply to the robustified FLC as well.

>>>>

\subsection{Open-Loop vs. Closed-Loop comparison}

\begin{figure}[t]
	\centering
	\includegraphics{Fig/ol_telemetries.pdf}
	\caption{Open-Loop method. Variations of longitudinal speed $\Delta u$, steering angle $\Delta\delta$, longitudinal force $\Delta F_x$, and lateral deviation $\Delta e$ relative to the nominal trajectory for three schemes --- TLC (track-limit), FLC (friction-limit), and TLC+FLC (both).
	Dashed lines mark the sign of the reference signal (positive, near-zero and negative).}

@@

\begin{figure}[t]
	\centering
	\includegraphics[scale = 1]{Fig/ol_saturation.pdf}
	\caption{Open-Loop. Axle saturation $S_j(\bx,\bu)$ for front axle (j=1, first panel) and rear axle (j=2, second panel) with respect to the curvilinear abscissa $\al$. Nominal configuration in dashed black, TLC in green, FLC in red, TLC+FLC in blue.
	%The dashed black line corresponds to the reference (nominal) configuration, while the green, red, and blue lines represent the robustified configurations TLC, FLC, and TLC+FLC, respectively.
	}
	\label{fig:ol_saturation}
\end{figure}

%It is of interest to analyze how a robustified constraint---or a combination of the two described in Sections~\ref{sec:FLC} and~\ref{sec:TLC}---influences the driving style.
We now investigate how the inclusion of different kinds of robustified constraints --- individually or in combination, as described in Sections~\ref{sec:FLC} and~\ref{sec:TLC} --- influences the driving style.
We compare the nominal feed-forward optimal trajectory, obtained without robustified constraints, with those resulting from a robustified track limit constraint (denoted as TLC), a robustified friction limit constraint (denoted as FLC), and a scenario where both constraints are robustified (denoted as TLC+FLC).

First we show results obtained with the open-loop method described in Section~\ref{sec:open_loop_planning}, setting $H=4$, and $\ga^\textrm{TLC}=\ga^\textrm{FLC}=1.28$, corresponding to a 90\% probability of satisfying both constraints.
%All the optimization are obtained with the open-loop method described in Section~\ref{sec:open_loop_planning}, setting $H=4$, and $\ga^\textrm{TLC}=\ga^\textrm{FLC}=1.28$, corresponding to a 90\% probability of satisfying both constraints.
%---Delta key signals
The comparison between the nominal case and the three robustified cases is illustrated in Figure~\ref{fig:ol_telemetries}, in terms of the deviation of selected signals from to their nominal counterparts.
The panels display variations on the longitudinal speed $\De u$ (first panel), wheel steering angle $\De \de$ (second panel), total longitudinal force $\De F_x$ (third panel), and lateral deviation of the center of mass (CoM) from the track centerline $\De e$ (fourth panel).
Except for the $\De u$ panel, each plot includes dashed black lines that indicate the sign of the nominal signal.
These indicators are shown at three distinct vertical levels --- upper, central, and lower --- corresponding respectively to positive, near-zero, and negative values of the nominal signal.

@@

This shift in trajectory and longitudinal speed becomes especially evident when approaching the sharp turn for $\al \in [0.707, 0.712]$. A noticeable increase in longitudinal speed is observed around $\alpha = 0.71$ for the TLC configuration  in the $\De u$ panel. This peak is followed by a sharp drop in speed, indicating intense braking.
The TLC+FLC configuration exhibits a qualitatively similar trend in $\Delta u$, although the peak is significantly lower and the subsequent speed reduction is clearly stronger. This behavior is consistent with the variation of the total longitudinal force $\Delta F_x$ shown in the third panel.
%\textcolor{red}{In particular, the TLC and TLC+FLC configurations exhibit a higher longitudinal force while approaching the sharp turn, in the curvilinear abscissa interval $\al\in\left[0.707, 0.712\right]$, resulting in a higher longitudinal speed.
%Subsequently, these two configurations apply a lower longitudinal force during braking compared to the nominal case, indicating more intense braking. This behavior explains the sharp reduction in longitudinal speed observed thereafter.}

@@

Firstly, we observe that the FLC and TLC+FLC configurations (red and blue lines) never reach full saturation ($S_j=1)$ due to the presence of the back-off term, which enforces a safety margin from the friction limit.
Secondly, we point out how the different trajectories followed by the TLC and TLC+FLC configurations (green and blue lines), compared to the nominal and FLC cases, entail distinct ground force demands, which are clearly visible here in terms of the $S_j$ ratio.
This behavior emerges around the sharp turn and becomes even more evident in the high-speed turn.
Specifically, in the interval $\al \in [0.745, 0.762]$, the TLC and TLC+FLC configurations begin to experience high axle loads approximately 12\,m before the nominal trajectory and sustain them for an additional 4.6\,m (TLC, green line) and 9.2\,m (TLC+FLC, blue line) beyond the nominal reference.

@@

\begin{figure}[t]
	\centering
	\includegraphics{Fig/cl_telemetries.pdf}
	\caption{Closed-Loop method. Variations of longitudinal speed $\De u$, steering angle $\De\de$, longitudinal force $\De F_x$, and lateral deviation $\De e$ relative to the nominal trajectory for three schemes --- TLC (track-limit), FLC (friction-limit), and TLC+FLC (both).
		Dashed lines mark the sign of the reference signal (positive, near-zero and negative).}

	%\caption{Evolution of the longitudinal speed $\De u$, wheel steering angle $\De \de$, longitudinal force $\De F_x$, and lateral deviation $\De e$ are reported for three robustification scenarios: TLC (Track Limit Constraints), FLC (Friction Limit Constraint), and TLC+FLC (both constraints enforced concurrently). Each profile is shown relative to the baseline (nominal) trajectory to highlight the effect of each constraint-handling strategy. The dashed lines in the last three panels indicate the sign of the reference signal, allowing one to determine whether a positive variation with respect to the baseline corresponds to an increase or a decrease in the absolute value. This distinction is particularly important for quantities that depend on the direction of the turn, such as the wheel steering angle $\de$ and the lateral deviation $e$, while in the case of $F_x$, it indicates whether the vehicle is accelerating or braking.}
	\label{fig:cl_telemetries}
\end{figure}

\begin{figure}[b!]
	\centering
	\includegraphics[scale = 1]{Fig/cl_saturation.pdf}
	\caption{Closed-Loop. Axle saturation $S_j(\bx,\bu)$ for front axle (j=1, first panel) and rear axle (j=2, second panel) with respect to the curvilinear abscissa $\al$. Nominal configuration in dashed black, TLC in green, FLC in red, TLC+FLC in blue.
		%The dashed black line corresponds to the reference (nominal) configuration, while the green, red, and blue lines represent the robustified configurations TLC, FLC, and TLC+FLC, respectively.
	}
	\label{fig:cl_saturation}
\end{figure}

An analogous analysis has been conducted using the closed-loop approach described in Section~\ref{sec:closed_loop_planning}, setting $\ga^\textrm{TLC}=\ga^\textrm{FLC}=1.28$.
The same key signals displayed for the previous case are shown in Figure~\ref{fig:cl_telemetries}, while the axle saturation ratios $S_j$ are illustrated in Figure~\ref{fig:cl_saturation}.
From one side, the results appear very similar to those obtained with the open-loop method. While the closed-loop method offers a more rigorous framework by enabling seamless covariance propagation over the entire lap, the observed agreement indicates that open-loop planning remains a valid and effective strategy, as long as the prediction horizon is appropriately selected.

On the other hand, the trajectories are also sufficiently different to appreciate some subtle but relevant distinctions. For instance, sector time increments w.r.t. the nominal case of corresponding configurations are lower than the ones arising from the open-loop method: 1.39\% for the TLC, 0.66\% for the FLC, and 2.07\% for the TLC+FLC.
Moreover, in both TLC and TLC+FLC configurations, the magnitude of $\De\de$ is appreciably smaller in the closed-loop case compared to the open-loop one, especially during the high-speed turn. 
Additionally, the TLC+FLC configuration no longer exhibits the throttle release in $\De F_x$ observed in the open-loop case.

As for the axle saturation, both methods lead to qualitatively similar evolutions, but slight differences can be observed within the high-speed turn interval $\al \in [0.745, 0.762]$. In this case, the TLC and TLC+FLC configurations begin to experience high axle loads approximately 10\,m before the nominal trajectory and sustain them for an additional 4.5\,m beyond the nominal reference. This is achieved through the closed-loop method's capability to \emph{steer} and tame the covariance matrix via the dynamic action of the controller.

%\subsection{Open-loop vs closed-loop robust planning}
Finally, we compare the back-off influence on the open-loop and closed-loop trajectories, using the robustified TLC with a higher $\ga^\textrm{TLC}=3.0$ for better visualization. As before, the open-loop planning is carried out with a prediction horizon of $H=4$ steps.
The back-off term $\be^{TLC}$ in~\eqref{eq:TLC} admits a clear geometrical interpretation: it corresponds to the amount by which the track boundaries are tightened to meet the probabilistic constraint, or, equivalently, to a virtual extension of the vehicle's width.

According to the latter interpretation, Figure~\ref{fig:olcl_traj_bands} shows two highlighted bands---red for the open-loop case and cyan for the closed-loop case---corresponding to the area swept by the back-off-extended-vehicle. Ultimately, the TLC enforces precisely that these bands remain entirely within the actual track limits. The comparison between the two regions illustrates how the additional conservativeness of the open-loop approach stems from larger back-off terms, i.e., a wider band that must stay within the track boundaries. The closed-loop method, on the other hand, benefits from the LQR-based feedback, which allows it to maintain smaller back-off terms and thus a narrower band, requiring a reduced safety margin from the track limits.
\begin{figure}
	\centering
	\includegraphics{Fig/olcl_traj_bands.pdf}
	\caption{Lorem ipsum.}
	\label{fig:olcl_traj_bands}
\end{figure}

\subsection{Simulations with noise realization}
To provide an empirical validation of the closed-loop robustified planning we simulated the evolution of the controlled vehicle with random initial conditions and included noise realizations. The random values for the initial state and the noise follow Gaussian distributions with covariance matrices $\bP_0$ and $\bQ$, respectively, as introduced in~\eqref{eq:dP}. This analysis cannot be performed for the open-loop approach, as it does not compute a stabilizing controller---resulting in trajectories that would significantly deviate from the track limits beyond the prediction horizon.
%For visual clarity, the comparison is limited to trajectories with the robustified TLC, but analogous considerations apply to the robustified FLC as well.

### 2025-06-06 stochastic_vehicle_dynamics.tex
[MM] New plot inserted

\begin{figure}
	\centering
	\input{Fig/Tikz_Sources/tikz_vehicle_model.tex}
	\caption{Single Track model. The ground force acting on each axle is decomposed into longitudinal and lateral components, indicated by red arrows. The longitudinal components are part of the input vector $\bu$, encompassing both acceleration and braking forces, while the lateral components are computed using Pacejka's Magic Formula based on the vertical load and slip angle at each axle. The figure also depicts the fixed reference frame and the body reference frame, with the origin located at the vehicle's center of mass (CoM). The state variables $u$ and $v$ represent the longitudinal and lateral velocities of the CoM w.r.t. the body frame, while $r$ denotes the yaw rate. The angle $\psi$ is used to describe the orientation of the body frame w.r.t. the fixed frame.}
	\label{fig:vehicle_model}
\end{figure}

>>>>

\begin{figure}
	\centering
	\input{Fig/Tikz_Sources/tikz_vehicle_model.tex}
	%\caption{Single Track model. The ground force acting on each axle is decomposed into longitudinal and lateral components, indicated by red arrows. The longitudinal components are part of the input vector $\bu$, encompassing both acceleration and braking forces, while the lateral components are computed using Pacejka's Magic Formula based on the vertical load and slip angle at each axle. The figure also depicts the fixed reference frame and the body reference frame, with the origin located at the vehicle's center of mass (CoM). The state variables $u$ and $v$ represent the longitudinal and lateral velocities of the CoM w.r.t. the body frame, while $r$ denotes the yaw rate. The angle $\psi$ is used to describe the orientation of the body frame w.r.t. the fixed frame.}
	\caption{Single Track model.}
	\label{fig:vehicle_model}
\end{figure}
